Kindly Bent to Free Us

### 2019-11-21 kindly/motivation.tex
Misc fixes for the iterator stuff.

\begin{lstlisting}
val fold :
  ('b:'k),('k <= aff_r) =>
  (&(aff_r1,'a) -> 'b -{aff_r}> 'b) -> &('a array) -> 'b -{lin}> 'b
\end{lstlisting}

The folded function receives a shared borrow of the element in the array.
The typing of the callback ensures
that this borrow is neither captured nor returned by the function.
%
This encapsulation is implemented with a universally quantified \emph{kind index variable} $r$.
The signature prescribes the type @&!(aff_r1,'a)@ for the
exclusive borrow of the resource with an affine kind at region nesting $r+1$. The return
type of the callback is constrained to kind @'k <= aff_r@
so that the callback certainly cannot return the borrowed argument.
%
The input of the fold is a shared borrow of the array,
which ensures that we have the right to share borrows of the inner content and
make multiple concurrent folds.

We can then easily use this function to fold over an array of files. Here
we compute the sum of their lengths:

\begin{lstlisting}
let all_lengths = fold (fun f s -> read f + s) &all_files 0 
\end{lstlisting}

@@

\begin{lstlisting}
val iter_mut : &!('a array) -> (&!(aff_r1,'a) -{un_r}> unit) -{lin}> unit
\end{lstlisting}

Here, the callback of the @iter@ function must be of kind @un_r@: Indeed, this
callback should be @un@, as it will be called multiple time, but it should also
be prevented from capturing objects of the region $r+1$ to prevent the borrow
from escaping.

While the distinction between mutable and immutable iteration function
might seem unfortunate, it is typical of
programming with borrows and is also present in the Rust standard library, for
instance. It allows the programmer to explicitly state how various iterations
are allowed to compose and be optimized.
\lang's region variables ensures that the content iterated on can never
be leaked outside of the iteration function.
This pattern is essential in many use-cases of linearity such
as pools of linear objects (see \cref{tuto:pool}).

>>>>

\begin{lstlisting}
val fold :
  ('b:'k),('k <= aff_r) =>
  (&(aff_r1,'a) -> 'b -{aff_r1}> 'b) -> &('k_1,'a array) -> 'b -{'k_1}> 'b
\end{lstlisting}

The folded function receives a shared borrow of the element in the array.
The typing of the callback ensures
that this borrow is neither captured nor returned by the function.
%
This encapsulation is implemented with a universally quantified \emph{kind index variable} $r$.
The signature prescribes the type @&(aff_r1,'a)@ for the
shared borrow of the resource with an affine kind at region nesting $r+1$. The return
type of the callback is constrained to kind @'k <= aff_r@
so that the callback certainly cannot return the borrowed argument.
%
The input of the fold is a shared borrow of the array,
which ensures that we have the right to share borrows of the inner content and
make multiple concurrent folds.

We can then easily use this function to fold over an array of files. Here
we compute the sum of their sizes:

\begin{lstlisting}
let total_size = fold (fun f s -> File.size f + s) &all_files 0 
\end{lstlisting}

@@

\begin{lstlisting}
val iter_mut : &!('a array) -> (&!(aff_r1,'a) -{un}> unit) -{aff}> unit
\end{lstlisting}

Since the callback will be called multiple times, it is explicitly
annotated with @un@.

While the distinction between mutable and immutable iteration function
might seem unfortunate, it is typical of
programming with borrows and is also present in the Rust standard library.
It allows the programmer to explicitly state how various iterations
are allowed to compose and be optimized.
\lang's region variables ensures that the content iterated on can never
be leaked outside of the iteration function.
This pattern is essential in many use-cases of linearity such
as pools of linear objects (see \cref{tuto:pool}).
